3.8: Polynomial Regression

### 2019-11-19 static/slides/20-slides.Rmd
Tweak slides 20

1. $X_1 \rightarrow Y$ is different for different levels of $X_1$
  - e.g. diminishing returns: $\uparrow X_1$ increases $Y$ at a *decreasing* rate
  - e.g. increasing returns: $\uparrow X_1$ increases $Y$ at an *increasing* rate

@@

# Polynomial Functions of $X$

$$\hat{Y_i} = \hat{\beta_0} + \hat{\beta_1} X_i + \hat{\beta_2} X_i^2 + \cdots + \hat{\beta_{\mathbf{r}}} X_i^{\mathbf{r}} + \epsilon_i$$

- Where $r$ is the highest power $X_i$ is raised to
  - e.g. quadratic $r=2$, cubic $r=3$, etc

@@

- Just another multivariate OLS regression model

@@

Marginal effect of GDP if GDP $=\$5$ (thousand):

@@

Marginal effect of GDP if GDP $=\$25$ (thousand):

@@

Marginal effect of GDP if GDP $=\$50$ (thousand):

@@

@@

>>>>

1. $X_1 \rightarrow Y$ is different for different levels of $X_1$
  - e.g. **diminishing returns**: $\uparrow X_1$ increases $Y$ at a *decreasing* rate
  - e.g. **increasing returns**: $\uparrow X_1$ increases $Y$ at an *increasing* rate

@@

# Polynomial Functions of $X$ I

- .blue[Linear], $X$

```{r, fig.retina=3}
library("mosaic")
linear=function(x){x}
quadratic=function(x){x^2}
cubic=function(x){x^3}
quartic=function(x){x^4}

ggplot(data.frame(x=c(-10,10)), aes(x=x))+
  stat_function(fun=linear, geom="line", size=2, color = "blue")+
    scale_x_continuous(breaks=seq(-1,1,1),
                     limits=c(-2,2),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(-1,1,1),
                     limits=c(-1,1))+
  #                   expand=expand_scale(mult=c(0,0.1)),
  #                   labels = function(x){paste("$", x, sep="")})+
  coord_cartesian(clip="off")+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)

```
]

# Polynomial Functions of $X$ I

- .blue[Linear], $X$

- .green[Quadratic], $X^2$

```{r, fig.retina=3}

ggplot(data.frame(x=c(-10,10)), aes(x=x))+
  stat_function(fun=linear, geom="line", size=2, color = "blue")+
  stat_function(fun=quadratic, geom="line", size=2, color = "green")+
    scale_x_continuous(breaks=seq(-1,1,1),
                     limits=c(-2,2),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(-1,1,1),
                     limits=c(-1,1))+
  #                   expand=expand_scale(mult=c(0,0.1)),
  #                   labels = function(x){paste("$", x, sep="")})+
  coord_cartesian(clip="off")+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)

```
]

---
# Polynomial Functions of $X$ I

- .blue[Linear], $X$

- .green[Quadratic], $X^2$

- .orange[Cubic], $X^3$

```{r, fig.retina=3}
ggplot(data.frame(x=c(-10,10)), aes(x=x))+
  stat_function(fun=linear, geom="line", size=2, color = "blue")+
  stat_function(fun=quadratic, geom="line", size=2, color = "green")+
  stat_function(fun=cubic, geom="line", size=2, color = "orange")+
    scale_x_continuous(breaks=seq(-1,1,1),
                     limits=c(-2,2),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(-1,1,1),
                     limits=c(-1,1))+
  #                   expand=expand_scale(mult=c(0,0.1)),
  #                   labels = function(x){paste("$", x, sep="")})+
  coord_cartesian(clip="off")+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)

```
]

# Polynomial Functions of $X$ I

- .blue[Linear], $X$

- .green[Quadratic], $X^2$

- .orange[Cubic], $X^3$

- .purple[Quartic], $X^4$
]

```{r, fig.retina=3}

ggplot(data.frame(x=c(-10,10)), aes(x=x))+
  stat_function(fun=linear, geom="line", size=2, color = "blue")+
  stat_function(fun=quadratic, geom="line", size=2, color = "green")+
  stat_function(fun=cubic, geom="line", size=2, color = "orange")+
  stat_function(fun=quartic, geom="line", size=2, color = "purple")+
    scale_x_continuous(breaks=seq(-1,1,1),
                     limits=c(-2,2),
                     expand=expand_scale(mult=c(0,0.1)))+
  scale_y_continuous(breaks=seq(-1,1,1),
                     limits=c(-1,1))+
  #                   expand=expand_scale(mult=c(0,0.1)),
  #                   labels = function(x){paste("$", x, sep="")})+
  coord_cartesian(clip="off")+
  labs(x = "X",
       y = "Y")+
  theme_classic(base_family = "Fira Sans Condensed", base_size=20)

```
]

# Polynomial Functions of $X$ I

$$\hat{Y_i} = \hat{\beta_0} + \hat{\beta_1} X_i + \hat{\beta_2} X_i^2 + \cdots + \hat{\beta_{\mathbf{r}}} X_i^{\mathbf{r}} + u_i$$
--

- Where $r$ is the highest power $X_i$ is raised to
  - quadratic $r=2$
  - cubic $r=3$

@@

- Just another multivariate OLS regression model!

class: inverse, center, middle

# The Quadratic Model

@@

Marginal effect of GDP if GDP $=5$ ($ thousand):

@@

Marginal effect of GDP if GDP $=25$ ($ thousand):

@@

Marginal effect of GDP if GDP $=50$ ($ thousand):

@@

class: inverse, center, middle

# The Quadratic Model: Maxima and Minima

@@

class: inverse, center, middle

# Are Polynomials Necessary?
